---
title: "Pycnogonide"
author: "Sébastien Rochette, ThinkR"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    includes:
      in_header: header_hide.html
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{03 - Pycnogonide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/javascript">
$(document).ready(function () {
  window.initializeCodeFolding("show" === "hide");
});
</script> 

# Exploration of "Pycnogonide" data

```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  out.width = "90%",
  fig.width = 7,
  fig.height = 6,
  fig.align = "center",
  eval = TRUE
  )
```

## Packages
```{r, message=FALSE}
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
# devtools::install_github("r-spatial/sf")
library(sf)
library(raster)
library(fasterize)
library(igraph)
library(rasterVis)
library(cowplot)
library(deeptools)
# devtools::document()
# devtools::document()
# devtools::load_all(here::here(""))
```

## Colours
```{r}
blue <- "#093564"
yellow <- "#efcc26"
grey <- "#675546"
```

## Load data
```{r}
# load data
export_file <- system.file("data_orig/export_last.csv", package = "deeptools")
liste_photo <- system.file("data_orig/liste_photo.txt", package = "deeptools")
```

## Prepare data
- Cleaning of species names to be easily usable
- Add user_id combining username and date of image analysis just in case a user sees the same image two times.

```{r}
mission2 <- readr::read_csv(export_file) %>%
  dplyr::select(-comment) %>%
  tidyr::extract(name,
          into = "datetime", regex = "_([[:digit:]]+).",
          remove = FALSE
  ) %>%
  mutate(datetime = ymd_hms(datetime, tz = "UTC")) %>%
  # clean names of species
  mutate(name_fr_clean = thinkr::clean_vec(name_fr, unique = FALSE)) %>% 
  group_by(username) %>%
  mutate(
    user_id = paste(username, as.character(as.numeric(as.factor(datDeb))), sep = "-")
  ) %>%
  ungroup()

# Separate observatory dataset
mission2_MAR <- mission2 %>% filter(obs_code == "MAR")
mission2_ONC <- mission2 %>% filter(obs_code == "JDF")
```

## Extraction of "pycnogonide"

```{r}
# Filter on Buccinide only
ONC2_pyc <- mission2_ONC %>% 
  filter(name_fr_clean == "pycnogonide")

# Filter and transform as spatial data
ONC2_pyc_carto <- mission2_ONC %>%
  to_carto(name_fr_clean, "pycnogonide")
```

## Exploration of annotations 
### Users

```{r}
ONC2_pyc %>% 
  count(username) %>% 
  arrange(desc(n))
```

### Images

- Number of annotations by image
```{r}
ONC2_pyc %>% 
  count(image_id) %>% 
  arrange(desc(n))
```

- Number of users by image
```{r}
ONC2_pyc %>% 
  group_by(image_id) %>% 
  summarize(n_users = length(unique(user_id))) %>% 
  arrange(desc(n_users))
```


### Example: `filter_col = username, filter_val = "grillus33", image_id == "10681"`

```{r, echo=FALSE}
ONC2_pyc %>% 
  filter(username == "grillus33") %>% 
  count(image_id) %>% 
  arrange(desc(n))

gg_users_image(x = ONC2_pyc_carto,
               filter_col = username, filter_val = "grillus33",
               image_id = 12738)

```

### Multiple users annotations

- Example with `image_id = 11425`

```{r}
gg_users_image(x = ONC2_pyc_carto,
               image_id = 11425)
```

- Comparison with buffer  
- Define a buffer size for future analyses 

> *It seems that precision is low for this point identification, then buffer size need to be big*

```{r}
gg_users_image(x = ONC2_pyc_carto,
               image_id = 11425, buffer = 40)
```

## Find all groups

As shown for "Buccinides", only one function is required to determine the differents groups of annotations in all images. This requires function `find_groups_in_all_images`, some time of computation and RAM available.


```{r, eval=FALSE}
# Chunk not evaluated in Rmd as results are saved
ONC2_pyc_carto_groups <- find_groups_in_all_images(ONC2_pyc_carto, .progress = TRUE, keep_list = FALSE, as_sf = FALSE, dist_buffer = 40)

if (!dir.exists(here::here("inst/outputs"))) {
  dir.create("inst/outputs", recursive = TRUE) 
}

readr::write_rds(
  ONC2_pyc_carto_groups,
  here::here("inst/outputs", "ONC2_pyc_carto_groups.rds"), 
  compress = "gz")

```
```{r}
outwd <- system.file("outputs", package = "deeptools")
ONC2_pyc_carto_groups <- readr::read_rds(file.path(outwd, "ONC2_pyc_carto_groups.rds"))
```


## Calculate statistics on images
*As a reminder, a group of objects is supposed to be a unique individual.*

### Number of objects per image (already known before)
```{r}
# Number of objects per image (already known before)
pyc_nobjects <- ONC2_pyc_carto_groups %>% 
  count(image_id) %>% 
  rename(n_objects = n) %>% 
  count(n_objects) %>% 
  arrange(desc(n)) %>% 
  rename(n_images = n)

# Number of marked objects by images
pyc_nobjects

# Plot
ggplot(pyc_nobjects) +
  geom_col(aes(x = n_objects, y = n_images), width = 1, 
           fill = yellow, colour = grey) +
  ggtitle("Total number of objects identified in images") +
  xlab("Number of objects by image") +
  ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```


### Statistics on groups

Calculate the number of users who marked the same individual. The `proportion` column can be used as a threshold to decide if the individual is kept. Indeed, an individual observed by only one of the users may not be a real individual.

```{r}
# Stats on groups
pyc_groups <- ONC2_pyc_carto_groups %>% 
  group_by(image_id, group_kept) %>% 
  summarise(
    n_users = n(),
    n_user_id = mean(n_user_id),
    proportion = n()/mean(n_user_id)
  ) %>% 
  ungroup()

# Number of individuals ~ proportion
pyc_groups

# Plot
pyc_groups %>% 
  ggplot() +
  geom_histogram(aes(proportion), bins = 10,
                 fill = yellow, colour = grey) +
  ggtitle("Number of individuals ~ proportion") +
  xlab("Proportion of users who identified an individual") +
  ylab("Number of individuals in this case") +
  theme(panel.background = element_rect(fill = blue))
```


### Statistics on number of groups by image

Calculate the number of groups in all images.

```{r}
## Stats on nb groups by image
pyc_ngroups_count <- pyc_groups %>% 
  group_by(image_id) %>% 
  summarise(n_groups = n()) 

# Number of images
pyc_ngroups_count %>% count(n_groups)

# Plot
ggplot(pyc_ngroups_count) +
  geom_histogram(aes(x = n_groups), bins = 30,
                 fill = yellow, colour = grey) +
  ggtitle("Number of groups (~individuals) in images") +
  xlab("Number of groups by image") + ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```


If we only keep groups identified by at least half of the users, we can recalculate the number of groups by image.

```{r}
## Stats on nb groups by image
pyc_ngroups_count_thd <- pyc_groups %>% 
  filter(proportion >= 0.5) %>% 
  group_by(image_id) %>% 
  summarise(n_groups = n()) 

# Number of images
pyc_ngroups_count_thd %>% count(n_groups)

# Plot
ggplot(pyc_ngroups_count_thd) +
  geom_histogram(aes(x = n_groups), bins = 30,
                 fill = yellow, colour = grey) +
  ggtitle("Number of groups (~individuals) in images") +
  xlab("Groups with proportion >= 0.5 by image") + ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```
