---
title: "Couverture de moules"
author: "Sébastien Rochette, ThinkR"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    includes:
      in_header: header_hide.html
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{02 - Couverture de moules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/javascript">
$(document).ready(function () {
  window.initializeCodeFolding("show" === "hide");
});
</script> 

# Exploration of "Couverture de moules" data

```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  out.width = "90%",
  fig.width = 7,
  fig.height = 6,
  fig.align = "center",
  eval = TRUE
  )
```


## Packages
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
# devtools::install_github("r-spatial/sf")
library(sf)
library(deeptools)
library(lubridate)
library(thinkr)
library(readr)
library(tidyr)
```

## Colours
```{r}
blue <- "#093564"
yellow <- "#efcc26"
grey <- "#675546"
```

## Load data
```{r}
# load data
export_file <- system.file("data_orig/export_last.csv", package = "deeptools")
liste_photo <- system.file("data_orig/liste_photo.txt", package = "deeptools")
```

## Prepare data
- Cleaning of species names to be easily usable
- Add user_id combining username and date of image analysis just in case a user sees the same image two times.

```{r}
mission2 <- read_csv(export_file) %>%
  dplyr::select(-comment) %>%
  tidyr::extract(name,
          into = "datetime", regex = "_([[:digit:]]+).",
          remove = FALSE
  ) %>%
  mutate(datetime = ymd_hms(datetime, tz = "UTC")) %>%
  # change coordinates as numeric
  mutate_at(vars(starts_with("pos"),
                 starts_with("middle"),
                 matches("length"),
                 starts_with("age")), as.numeric) %>%
  # clean names of species
  mutate(name_fr_clean = clean_vec(name_fr, unique = FALSE)) %>% 
  group_by(username) %>%
  mutate(
    user_id = paste(username, as.character(as.numeric(as.factor(datDeb))), sep = "-")
  ) %>%
  ungroup()

# Separate observatory dataset
mission2_MAR <- mission2 %>% filter(obs_code == "MAR")
mission2_ONC <- mission2 %>% filter(obs_code == "JDF")
```

## Extraction of "Buccinide"

Function `to_carto` extract and transform data as spatial object for following analyses.

```{r}
# Filter on Buccinide only
MAR_mussel <- mission2_MAR %>% 
  filter(name_fr_clean == "couverture_de_moules")

# Filter and transform as spatial data
MAR_mussel_carto <- mission2_MAR %>%
  filter(name_fr_clean == "couverture_de_moules") %>% 
  to_carto(name_fr_clean, "couverture_de_moules")
```

_Only because it is nice_
```{r}
ggplot(MAR_mussel_carto) +
  geom_sf(aes(fill = as.character(image_id)), alpha = 0.1) +
  guides(fill = FALSE)
```

## Exploration of annotations 
### Users

```{r}
MAR_mussel %>% 
  count(username) %>% 
  arrange(desc(n))
```

### Images

- Number of annotations by image
```{r}
MAR_mussel %>% 
  count(image_id) %>% 
  arrange(desc(n))
```

- Number of users by image
```{r}
MAR_mussel %>% 
  group_by(image_id) %>% 
  summarize(n_users = length(unique(user_id))) %>% 
  arrange(desc(n_users))
```


### Example: `filter_col = username, filter_val = "grillus33", image_id == "14190"`

```{r, echo=FALSE}
MAR_mussel %>% 
  filter(username == "grillus33") %>% 
  count(image_id) %>% 
  arrange(desc(n))

gg_users_image(x = MAR_mussel_carto,
               filter_col = username, filter_val = "grillus33",
               image_id = 14190)

```


### Multiple users annotations

- Example with `image_id = 14190`

```{r}
gg_users_image(x = MAR_mussel_carto,
               image_id = 14190)
```

- Because we work with polygons, buffer is not totally necessary. However, just in case there are small areas, we can set a small buffer.

```{r}
gg_users_image(x = MAR_mussel_carto,
               image_id = 14190, buffer = 5)
```

## Find all groups

As shown for "Buccinides", only one function is required to determine the differents groups of annotations in all images. This requires function `find_groups_in_all_images`, some time of computation and RAM available.


```{r, eval=FALSE}
# Chunk not evaluated in Rmd as results are saved
MAR_mussel_carto_groups <- MAR_mussel_carto %>% 
  find_groups_in_all_images(.progress = TRUE, keep_list = FALSE,
                            as_sf = FALSE, dist_buffer = 5)

if (!dir.exists(here::here("inst/outputs"))) {
  dir.create("inst/outputs", recursive = TRUE) 
}

readr::write_rds(
  MAR_mussel_carto_groups,
  here::here("inst/outputs", "MAR_mussel_carto_groups.rds"), 
  compress = "gz")

```
```{r}
outwd <- system.file("outputs", package = "deeptools")
MAR_mussel_carto_groups <- readr::read_rds(file.path(outwd, "MAR_mussel_carto_groups.rds"))
```


## Calculate statistics on images
*As a reminder, a group of objects is supposed to be a unique individual.*

### Number of objects per image (already known before)
```{r}
# Number of objects per image (already known before)
mussel_nobjects <- MAR_mussel_carto_groups %>% 
  count(image_id) %>% 
  rename(n_objects = n) %>% 
  count(n_objects) %>% 
  arrange(desc(n)) %>% 
  rename(n_images = n)

# Number of marked objects by images
mussel_nobjects

# Plot
ggplot(mussel_nobjects) +
  geom_col(aes(x = n_objects, y = n_images), width = 1, 
           fill = yellow, colour = grey) +
  ggtitle("Total number of objects identified in images") +
  xlab("Number of objects by image") +
  ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```

### Calculate surfaces of polygons

- We calculte the average and the standard deviation of groups

```{r}
MAR_mussel_carto_groups_area <- MAR_mussel_carto_groups %>% 
  mutate(area = MAR_mussel_carto_groups %>% 
           st_sf() %>% st_area()) %>% 
  group_by(image_id, group_kept) %>% 
  summarise(nb_polygons_in_group = n(),
            area_mean = mean(area),
            area_sd = sd(area),
            area_sd = if_else(is.na(area_sd), 0, area_sd),
            area_cv = area_sd / area_mean) %>% 
  arrange(desc(nb_polygons_in_group))

MAR_mussel_carto_groups_area
```

- Graph of areas

```{r}
MAR_mussel_carto_groups_area %>% 
  ggplot() +
  geom_histogram(aes(area_mean), bins = 50,
             fill = yellow, colour = grey) +
  ggtitle("Average areas of groups identified in images") +
  xlab("Area (log scale)") +
  ylab("Number of groups") +
  theme(panel.background = element_rect(fill = blue)) +
  scale_x_log10()
```

