---
title: "Buccinides"
author: "Sébastien Rochette, ThinkR"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    includes:
      in_header: header_hide.html
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{02 - Buccinides}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/javascript">
$(document).ready(function () {
  window.initializeCodeFolding("show" === "hide");
});
</script> 

# Exploration of "Buccinide" data

```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  out.width = "90%",
  fig.width = 7,
  fig.height = 6,
  fig.align = "center",
  eval = TRUE
  )
```
  
## Packages
```{r, message=FALSE}
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
# devtools::install_github("r-spatial/sf")
library(sf)
library(raster)
library(fasterize)
library(igraph)
library(rasterVis)
library(cowplot)
library(deeptools)
# devtools::document()
# devtools::document()
# devtools::load_all(here::here(""))
```

## Colours
```{r}
blue <- "#093564"
yellow <- "#efcc26"
grey <- "#675546"
```

## Load data
```{r}
# load data
export_file <- system.file("data_orig/export_last.csv", package = "deeptools")
liste_photo <- system.file("data_orig/liste_photo.txt", package = "deeptools")
```

## Prepare data
- Cleaning of species names to be easily usable
- Add user_id combining username and date of image analysis just in case a user sees the same image two times.

```{r}
mission2 <- readr::read_csv(export_file) %>%
  dplyr::select(-comment) %>%
  tidyr::extract(name,
          into = "datetime", regex = "_([[:digit:]]+).",
          remove = FALSE
  ) %>%
  mutate(datetime = ymd_hms(datetime, tz = "UTC")) %>%
  # clean names of species
  mutate(name_fr_clean = thinkr::clean_vec(name_fr, unique = FALSE)) %>% 
  group_by(username) %>%
  mutate(
    user_id = paste(username, as.character(as.numeric(as.factor(datDeb))), sep = "-")
  ) %>%
  ungroup()

# Separate observatory dataset
mission2_MAR <- mission2 %>% filter(obs_code == "MAR")
mission2_ONC <- mission2 %>% filter(obs_code == "JDF")
```

## Extraction of "Buccinide"

Function `to_carto` extract and transform data as spatial object for following analyses.

```{r}
# Filter on Buccinide only
ONC2_bucc <- mission2_ONC %>% 
  filter(name_fr_clean == "escargot_buccinide")

# Filter and transform as spatial data
ONC2_bucc_carto <- mission2_ONC %>%
  to_carto(name_fr_clean, "escargot_buccinide")
```

## Exploration of annotations 
### Users

```{r}
ONC2_bucc %>% 
  count(username) %>% 
  arrange(desc(n))
```

### Images

```{r}
ONC2_bucc %>% 
  count(image_id) %>% 
  arrange(desc(n))
```

### Example: `filter_col = username, filter_val = "Lawzama", image_id == "10681"`

```{r, echo=FALSE}
ONC2_bucc %>% 
  filter(username == "Lawzama") %>% 
  count(image_id) %>% 
  arrange(desc(n))

gg_users_image(x = ONC2_bucc_carto,
               filter_col = username, filter_val = "Lawzama",
               image_id = 10681)

```

### Example: `filter_col = username, filter_val = "Kazu", image_id == "10681"`

```{r}
gg_users_image(x = ONC2_bucc_carto,
               filter_col = username, filter_val = "Kazu",
               image_id = 10681)
```

### Example: `filter_col = username, filter_val = c("Kazu", "Lawzama"), image_id == "10681"`

```{r}
gg_users_image(x = ONC2_bucc_carto,
               filter_col = username, filter_val = c("Kazu", "Lawzama"),
               image_id = 10681)
```

## Compare annotations of the same image
### Super-impose annotations
This allows to check visually what sizes of buffer we can use to determine animals discovered by two persons.

```{r}
# Number of animals seen on the image by username
ONC2_bucc %>% 
  filter(image_id == 10681) %>% 
  count(user_id)

gg_users_image(x = ONC2_bucc_carto,
               image_id = 10681, buffer = 10)  

```

### Find common animals between users

#### Method detailed

*These steps do not have to be run as they are included in `find_groups_in_image` function presented below. These steps are here to explain the method.*

To compute statistics on individuals on images, it is important to know if different users found the same individuals. There is a variability in positioning individuals. Hence, to find out common individuals between users, we need to look in the neighborhood.
A quick exploration suggests a buffer area of 10 pixels could be enough for finding overlapping annotations.  
The following steps are here to detail the complete process, but at the end, they are all included in a unique fonction `find_commons`. *Note that functions are adapted to the structure of the examle dataset, included column names.*

For each image steps are:

- Create Voronoi polygons around sf features by `user_id`. Voronoi is necessary to get non overlapping buffer areas of too close individuals. This also allows for super-imposed individuals.
- Crop Voronoi with buffer area. This avoids to look in a too far neighborhood.
- Transform Voronoi as raster by `user_id`. Transforming polygons into pixels allow to find out the most overlapping groups of polygons among users.
- Stack all rasters by `image_id`

```{r, fig.width=10, fig.height=4, out.width='90%'}
# Choose one image for one user
image_id <- "10681"
user_id <- "Kazu-100"

bucc_image_user <- ONC2_bucc_carto %>% 
  dplyr::filter(image_id == !!image_id,
                user_id == !!user_id)

r_image <- raster(bucc_image_user, res = 0.5)
dist_buffer <- 10

# Calculate voronoi for one user_id
image_user_intermediates <- voronoi_rasterize(
  x = bucc_image_user,
  dist_buffer = dist_buffer,
  r_image = r_image,
  keep_intermediate = TRUE)

# Show intermediate steps
p1 <- ggplot(image_user_intermediates$points_in_voronoi) +
  geom_sf(aes(fill = image_pol_id)) +
  geom_sf(data = st_cast(st_geometry(bucc_image_user), "POINT"), size = 0.25) +
  theme_images(x = image_user_intermediates$points_in_voronoi, fill = "c", color = NULL)
  
p2 <- ggplot(image_user_intermediates$points_in_voronoi_in_buffer) +
  geom_sf(aes(fill = image_pol_id)) +
    geom_sf(data = st_cast(st_geometry(bucc_image_user), "POINT"), size = 0.25) +
  theme_images(x = image_user_intermediates$points_in_voronoi, fill = "c", color = NULL)

p3 <- gplot(image_user_intermediates$voronoi_in_buffer_as_raster) +
  geom_tile(aes(fill = value)) +
  theme_images(x = image_user_intermediates$points_in_voronoi, fill = "c", color = NULL, na.value = NA)

cowplot::plot_grid(plotlist = list(p1, p2, p3), ncol = 3)

```

The above intermediate steps are included in the `voronoi_stacker` function, which stacks rasters of all `user_id`.

```{r}
image_id <- "10681"
bucc_image <- ONC2_bucc_carto %>% 
  dplyr::filter(image_id == !!image_id)

r_image <- raster(bucc_image, res = 0.5)
dist_buffer <- 15

bucc_voronoi_stack <- voronoi_stacker(x = bucc_image,
           dist_buffer = dist_buffer,
           r_image = r_image)
  
gplot(bucc_voronoi_stack) +
  geom_tile(aes(fill = value)) +
  facet_wrap(~variable) +
  theme_images(x = image_user_intermediates$points_in_voronoi,
               fill = "c", color = NULL, na.value = NA) +
  ggtitle("Pixels numbered according to marked individuals")
```

For each pixel of the stack, we can define a group of polygons with the same size as the number of `user_id`. Indeed, an important part of pixels are not covered by polygons and will be identified as `NA-...-NA-...-NA`. Once removed those pixels, we can find the most represented groups of polygons.  

```{r}
# Combine layers and find groups of polygons
bucc_groups_count <- group_pixels_count(bucc_voronoi_stack)
bucc_groups_count
```

Problem is that some polygons are in more or less big groups, sometimes being in combination with two different polygons of the same user. We need to find out the best combinations of polygons to associate them to individuals really appearing on the original images.  
In step 1, for each polygon independently:

- Choose group when associated to maximum other polygons
- Choose group with highest surface in common

With these two rules, some chosen groups may include polygons associated to other groups.

```{r}
#' Find in which groups are each polygons
all_image_pol_ids <- pull(bucc_image, image_pol_id) %>% unique()

bucc_groups_top <- find_top_groups(bucc_groups_count,
                                   all_image_pol_ids)
test_ids_in_group <- test_groups_kept(bucc_groups_top)

# Show in how many groups are individuals (Should be only one)
test_ids_in_group$ids_in_groups_count
```

It is necessary to create a loop to reduce grouping possibilities based on same rules.  
In step two, for each `image_pol_id` found in multiple groups, keep the only group following the above two rules. Others are removed from list of possibilities. Step one is run again without these groups. Run these two steps until each polygon is found in one and only one group. *If an `image_pol_id` has no group left, it is included in a group alone*

```{r}
# If image_pol_ids are not unique reduce possibilities until it is good
if (test_ids_in_group$max_groups > 1) {
  # Run again group selection while removing groups with problems
  # _Find groups to remove
  test_ids_in_group2 <- test_ids_in_group
  bucc_groups_top2 <- bucc_groups_top
  group_remove2 <- NULL
  
  while (test_ids_in_group2$max_groups > 1) {

    group_remove <- test_ids_in_group2$ids_in_groups_count %>% 
      filter(n > 1) %>% 
      inner_join(test_ids_in_group2$ids_in_groups, by = "list_ids") %>% 
      inner_join(bucc_groups_top2, by = "group_kept") %>% 
      dplyr::select(-image_pol_id) %>% 
      distinct() %>% 
      group_by(list_ids) %>% 
      arrange(desc(n_pols), desc(n_pixels)) %>% 
      slice(-1) %>% 
      pull(group_kept) %>% 
      unique()
    
    group_remove2 <- unique(c(group_remove2, group_remove))
    
    bucc_groups_top2 <- bucc_groups_count %>% 
      filter(!grouped_ids %in% group_remove2) %>% 
      find_top_groups(all_ids = all_image_pol_ids)
    
    # bucc_groups_top2 %>% filter(grepl(7, group_kept))
    # bucc_groups_top2 %>% filter(grepl(32, group_kept))
    
    test_ids_in_group2 <- test_groups_kept(bucc_groups_top2)
    
  }
  # Retrieve success grouping
  bucc_groups_top <- bucc_groups_top2
  # Last verification
  test_ids_in_group <- test_groups_kept(bucc_groups_top)
}

# Show in how many groups are individuals (Should be only one)
test_ids_in_group$ids_in_groups_count
```

#### Function: Find groups in one image

The above procedure is included in a unique function available for a unique image: `find_groups_in_image(x, image_id)` which adds the group name to the original dataset.  
We can then verify the grouping procedure visually.

```{r}
# Add group names in image_sf
bucc_image_grouped <- ONC2_bucc_carto %>% 
  find_groups_in_image(image_id = "10681")

# Create specific image with group names
bucc_image_grouped_groups <- bucc_image_grouped %>% 
  group_by(group_kept) %>% 
  summarize() %>% 
  st_centroid() %>% 
  cbind(st_coordinates(.))

ggplot(bucc_image_grouped %>% 
         mutate(group_kept = 
         forcats::fct_reorder(group_kept, desc(n_pols)))
       ) +
    geom_sf(aes(color = group_kept),
      show.legend = "line",
      size = 2 #alpha = 0.1
    ) +
  ggrepel::geom_text_repel(
    data = bucc_image_grouped_groups,
    aes(x = X, y = Y, label = group_kept)) +
    theme_images(x = bucc_image_grouped, fill = NULL, color = "d", na.value = "grey20") +
  guides(color = FALSE) +
  ggtitle("Identification of groups of marked individuals")
```

### Find all groups for all images

Everything can be included in a unique function `find_groups_in_all_images` to explore the entire dataset at once.  
This takes some time and some place:  

- ~30min on dual core
- Be sure to have at least **10Go RAM** available, otherwise use `find_groups_in_image` iteratively in a loop for instance.

```{r, eval=FALSE}
# Chunk not evaluated in Rmd as results are saved
ONC2_bucc_carto_groups <- find_groups_in_all_images(ONC2_bucc_carto, .progress = TRUE, keep_list = FALSE, as_sf = FALSE)

if (!dir.exists(here::here("inst/outputs"))) {
 dir.create("inst/outputs", recursive = TRUE) 
}

readr::write_rds(
  ONC2_bucc_carto_groups,
  here::here("inst/outputs", "ONC2_bucc_carto_groups.rds"), 
  compress = "gz")

```
```{r}
outwd <- system.file("outputs", package = "deeptools")
ONC2_bucc_carto_groups <- readr::read_rds(file.path(outwd, "ONC2_bucc_carto_groups.rds"))
```

## Calculate statistics on images
*As a reminder, a group of objects is supposed to be a unique individual.*

### Number of objects per image (already known before)
```{r}
# ONC2_bucc_carto_groups 

# Number of objects per image (already known before)
bucc_nobjects <- ONC2_bucc_carto_groups %>% 
  count(image_id) %>% 
  rename(n_objects = n) %>% 
  count(n_objects) %>% 
  arrange(desc(n)) %>% 
  rename(n_images = n)

# Number of marked objects by images
bucc_nobjects

# Plot
ggplot(bucc_nobjects) +
  geom_col(aes(x = n_objects, y = n_images), width = 1, 
           fill = yellow, colour = grey) +
  ggtitle("Total number of objects identified in images") +
  xlab("Number of objects by image") +
  ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```

### Statistics on groups

Calculate the number of users who marked the same individual. The `proportion` column can be used as a threshold to decide if the individual is kept. Indeed, an individual observed by only one of the users may not be a real individual.

```{r}
# Stats on groups
bucc_groups <- ONC2_bucc_carto_groups %>% 
  group_by(image_id, group_kept) %>% 
  summarise(
    n_users = n(),
    n_user_id = mean(n_user_id),
    proportion = n()/mean(n_user_id)
  ) %>% 
  ungroup()

# Number of individuals ~ proportion
bucc_groups

# Plot
bucc_groups %>% 
  ggplot() +
  geom_histogram(aes(proportion), bins = 10,
                 fill = yellow, colour = grey) +
  ggtitle("Number of individuals ~ proportion") +
  xlab("Proportion of users who identified an individual") +
  ylab("Number of individuals in this case") +
  theme(panel.background = element_rect(fill = blue))
```

### Statistics on number of groups by image

Calculate the number of groups in all images.

```{r}
## Stats on nb groups by image
bucc_ngroups_count <- bucc_groups %>% 
  group_by(image_id) %>% 
  summarise(n_groups = n()) 

# Number of images
bucc_ngroups_count %>% count(n_groups)

# Plot
ggplot(bucc_ngroups_count) +
  geom_histogram(aes(x = n_groups), bins = 30,
                 fill = yellow, colour = grey) +
  ggtitle("Number of groups (~individuals) in images") +
  xlab("Number of groups by image") + ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```

If we only keep groups identified by at least half of the users, we can recalculate the number of groups by image.

```{r}
## Stats on nb groups by image
bucc_ngroups_count_thd <- bucc_groups %>% 
  filter(proportion >= 0.5) %>% 
  group_by(image_id) %>% 
  summarise(n_groups = n()) 

# Number of images
bucc_ngroups_count_thd %>% count(n_groups)

# Plot
ggplot(bucc_ngroups_count_thd) +
  geom_histogram(aes(x = n_groups), bins = 30,
                 fill = yellow, colour = grey) +
  ggtitle("Number of groups (~individuals) in images") +
  xlab("Groups with proportion >= 0.5 by image") + ylab("Number of images") +
  theme(panel.background = element_rect(fill = blue))
```

## Estimate average size of individuals

*We assume that a group is a unique individual.*  
We also remove groups identified by less than 50% of users.

```{r}
# Mean size of individuals
bucc_lengths <- ONC2_bucc_carto_groups %>% 
  left_join(bucc_groups %>% 
              dplyr::select(image_id, group_kept, proportion),
            by = c("image_id", "group_kept")) %>% 
  filter(proportion >= 0.5) %>% 
  group_by(image_id, group_kept) %>% 
  summarise(mean_length = mean(length))

ggplot(bucc_lengths) +
  geom_histogram(aes(mean_length), bins = 50,
                 fill = yellow, colour = grey) +
  scale_x_log10() +
  theme(panel.background = element_rect(fill = blue)) +
  ggtitle("Distribution of lengths of individuals") +
  xlab("Group mean length (log scale)") +
  ylab("Number of individuals")

```
